\documentclass[nofootbib,floatfix,11pt]{article}
\usepackage[square]{natbib}
\usepackage[paperwidth=8.5in,paperheight=11in,centering,margin=1in]{geometry}
\usepackage{paralist}
\usepackage{parskip}
\parskip=5pt

% suck up extra white space
%\setlength{\parskip}{3pt}
\setlength{\parsep}{0pt}
\setlength{\headsep}{0pt}
\setlength{\topskip}{0pt}
\setlength{\topmargin}{0pt}
\setlength{\topsep}{0pt}
\setlength{\partopsep}{0pt}
\usepackage{tabularx}

\renewcommand{\thefootnote}{\alph{footnote}}
\setlength{\belowcaptionskip}{-10pt}

\usepackage{amsmath}
\usepackage{amsbsy}

\pagestyle{empty}

\begin{document} 

\begin{center}
{\Large The  LSST Data Science Fellowship Program}

Director: Lucianne Walkowicz (Adler Planetarium and University of Chicago)\\
Andy Connolly (University of Washington),  Zeljko Ivezic (University of Washington), Mario Juric  (University of Washington), Vicky Kalogera (Northwestern University), Chris Lintott (Oxford University), Robert Lupton (Princeton University), Phil Marshall (SLAC National Lab and Stanford University).

\end{center}

\bigskip

\section{Introduction}

With a new generation of ground and space-based surveys generating petabytes of data per year, together with the increasing complexity of data coming from these instruments, astrophysics is becoming ever more dependent on developments in computing. There is, however, a disconnect between the skills that are needed for an era rich in data and those that we teach to incoming graduate students and early-career postdocs. We propose here a program that will enhance a traditional astrophysics curriculum; extending a strong physics education to one that encompasses computational techniques, programming skills, data management, statistics, and data analysis.

There are a number of exemplars in the astronomy community for how to train cohorts of students to address a common set of science questions. One of the most successful of these was the European Union sponsored Gaia Research for European Astronomy Training (GREAT) Network. The goal of the GREAT network was to equip students ``with the skills and expertise to become future leaders in astronomy or to enter industry, especially in the area of information and communication technologies''.
Over a period of $3-5$ years, they trained cohorts of students through a series of schools and workshops with a coherent curriculum that addressed both the science and technical aspects of working with Gaia data. The impact of this program was not just a set of students with the right skills but also a community, working in a common field, who knew and interacted with one another throughout the year. 

This proposal describes such a program tailored to the science of all of the LSST science collaborations; the LSST Science Fellowship Program. 


\section{An ongoing training program}

The primary goals of the LSST Data Science Fellowship program are to teach the skills required for LSST science that are not easily addressed by current astrophysics graduate programs (i.e.\ to enhance and not replace the current physics based curricula). By creating a two-year training program (with three one-week schools per year) we expect to have sufficient time to teach the appropriate skills, to interleave hands-on applications using precursor data so that the training will be directly applicable to a student's research, and to create an online resource that could be used by the broader astronomical community (i.e.\ a curriculum with associated teaching material).  We propose to start this program in summer of 2016 with an initial cohort of 15 students and bring in (and graduate) 15 new students per year (once we complete the first two-year program, so from year 2 and on). A 2016 start will provide sufficient time to complete the development of the curriculum and logistics so that each workshop or school will build upon the material presented in earlier schools.

With a recycle time of two years (50\% of the cohort entering and leaving each year) there is the opportunity for rolling cohorts, which means that we can implement peer based learning schemes such as pair or quad coding, which have been shown to be highly effective for the acquisition of skills. By keeping the cohort small we enable not just the training of students but also the development of a community, working in common aspects of data, who know an interact with one another on a regular basis and will continue to interact in the time between workshops.  The curriculum will adopt a spiral philosophy where all aspects of the topics described below will be introduced in each workshop (building upon previous material). Each workshop (lasting a week) will build upon the material taught in the previous workshop. By distributing the workshops throughout the year (as opposed to a single three-week workshop in the summer) we will ensure that the skills are continuously reinforced and that the student community becomes a coherent body who know one another and can interact with one another in the period between workshops. 



Each workshop may emphasize a particular theme but statistics, data management, software engineering, visualization will be incorporated into all of the workshops. Besides lectures, the workshops will emphasize group projects so that the majority of the work is hands-on. These hands-on components will focus on the primary science objectives drawn from all of the LSSTC science collaborations including cadence optimization, software and data for level-3 products, transient classification, and moving sources.

Students who do not yet possess basic proficiency in python will be expected to do some homework prior to the workshop so they start to bring themselves up to speed (e.g. courses available from learnpythonthehardway, coursera, codeacademy, etc.). Our goal is not to teach basic syntax of coding so much as a conceptual framework and best practices for code development and software engineering. 

%Students will be coached on collaborative processes and how to educate each other both at the workshop and within the community beyond.

We envisage an initial program that will run for 3 years and educate and graduate two cohorts of students (a minimum of 30 students, but possibly up to 60 students, see budget section). This will provide sufficient time to evaluate the impact of the program. The timeline (starting in the summer of  2016) is set to enable these cohorts to have graduated from the program by the start of LSST commissioning. Assuming that we enroll second year graduate students into the program (i.e.\ those who have completed much of their initial physics prerequisites), the first set of graduate students will also be graduating and becoming postdocs at the time of commissioning. This will seed both the LSSTC and broader astronomy community with expert researchers and educational ambassadors for these kinds of LSST-relevant skills.


The faculty for the program will be drawn primarily from LSSTC members who are not directly involved in the construction of the project. Project personnel may be lecturers for the summer schools where project specific knowledge needs to be imparted to the students. The program will be led by Lucianne Walkowicz (Adler Planetarium and University of Chicago). Lucianne will serve as the director of the school for the first three years of the program. Members of the LSSTC who have expressed an interest in teaching at the school include: Sarah Bridle, Tamas Budavari, Andy Connolly, Dan Foreman-MacKey, Zeljko Ivezic, Mario Juric, Vicky Kalogera, Chris Lintott, Robert Lupton, Phil Marshall, Bryce Menard, Maria Suveges, Jake vanderPlas, Michael Wood-Vasey


A proposed curriculum for these summer schools would include the following areas: 

\begin{itemize}
\item “Data Science Basics”: an introduction to statistics, machine learning, and information theory including a brief introduction to common tools used in software engineering (e.g. code repositories, issue tracking, GitHub)
\item Image processing: an introduction to and hands-on experience with the LSST image processing tools including the analysis of existing data sets
\item Non-supervised machine learning: the application and use of density estimation and clustering techniques with astronomical data
\item Supervised machine learning and time series: classification of multispectral data and time series analysis of variable sources with incomplete and noisy sampling
\item Scalable programing and data management: use of databases and parallel programming that can enable analyses to scale to large data sets
\item Data visualization: visualization of LSST images and catalog data including interactive visualization
\end{itemize}

Schools will be rotated through the LSSTC member institutes to maximize the exposure of the students to the research that is ongoing throughout the consortium and to enable LSSTC institutions to advertise themselves to an up and coming (and well trained) set of grad students and postdocs. By making use of this diversity of skills within our collaboration we can seed our institutes with grad students and postdocs who have been trained and who can then train others within their own individual universities (an infection model). To minimize the cost of the program any institute who will host one of the summer schools will be required to commit to provide postdocs and advanced graduate students who will TA the hands-on activities. This will help engage the postdoctoral communities within the LSSTC within this school. 

All materials generated for this program will be made available as open source and will be accessible to all students within and beyond the LSSTC member institutes.

All students who complete the program and graduate will be awarded the title of LSST Data Science Fellow to demonstrate that they have attended the school.

\section{Engaging the community and deliverables}

The program will be advertised to all LSSTC institutions, the science collaborations, and the general scientific community (including at the science collaborations all hands meetings). All LSSTC member institutions will be offered the opportunity to host one of the workshops with the selection of workshop hosts chosen to ensure a diverse representation of the LSSTC community (both geographically and scientifically). Student selection (and lecturer selection) will be undertaken by an advisory panel drawn from the LSSTC to ensure a diverse population (with an emphasis on under-represented groups) .

The deliverables for this program, aside from a trained graduate community for the LSSTC, will included all the educational material generated through the school. This will be in the form of teaching material, videos of the classes and lectures, and hands on exercises. These materials will be available in a single portal accessible over the web. We anticipate that these materials will be used by LSST fellows to reinforce what they have learned at the schools, will provide a mechanism by which non-LSST fellows can follow the program, and will be used by other institutes and universities to supplement their own educational programs and classes. 

We expect that the TA's (postdocs and senior graduate students) of the universities hosting the schools will benefit substantially from exposure to the lectures and educational material and from interactions with the lecturers and students.

\section{Budget}

For this program to be developed properly it is important that one person takes major responsibility for the development and organization of all aspects of the program. We envision that a postdoctoral fellow with experience in LSST science and data methodology should be recruited who will spend at least 50\% of their time on the proposed training program. In order to attract a person with high level of qualifications, it is necessary to be able to offer a full time position, so the person can spend up to 50\% of their time on LSST-related research. This postdoc will support the director, develop curricula for the program, manage the web-based material, interact with the teaching faculty to ensure that the teaching material is consistent with the curriculum, develop hands on activities for the schools, and work with the institutional TAs to ensure that they are trained in the tools used by the program.

We provide three funding scenarios for the program. These scenarios are: (i) the full-funding model, where LSSTC funds the entirety of the  program, (ii) the Northwestern model, where Northwestern (through CIERA) provides 50\% of the lead postdoctoral fellow and this support triggers an equal amount of matching funds from the Gates-Simonyi pledge, and the rest of the funds are provided from the LSSTC, and (iii) the minimal model, like the Northwestern model, but LSSTC provides funds for only 50\% of student travel support. 

For all scenarios we propose an initial 2 day workshop in 2015 for the program instructors to discuss and define the structure of the curriculum and materials that will need to be developed. Support for this is requested at a level of \$5K in Year1.

The cost of a postdoctoral fellow (at a salary of \$55K and using the average Northwestern 28.2\% benefits rate and no overhead) would be \$70.5K per year. Travel expenses for a student to attend three one-week workshops/schools per year is estimated at \$4,000 per student. Support for faculty who teach at the schools is \$1K for travel per faculty per school or \$9K per year (assuming three faculty per school).

We assume that expense reimbursements will be handled through the LSSTC (including student travel and local expenses for food etc) or by the Northwestern CIERA staff to minimize administrative costs.

We note that additional students (up to an additional 15 per year) could be accommodated in the proposed program, if they paid their own costs. In this case, by the end of the three-year period we will be able to graduate a total of up to 60 students. 

We will seek additional support from foundations, the NSF and DOE for this program to increase the size of the graduate student cohort and to provide a long-term basis for the program.

% We have initiated a discussion with Earnestine Easter a program officer in NSF's Division of Graduate  Education  to %use this workshop as a model for a new way of teaching graduate students. {\bf I am not sure that we should include this statement; why are these funds needed? If we get these funds, does it mean we %don't need LSSTC funds? If we get them in addition to the LSSTC funds, what more would we do? Are these funds envisioned for growing the number of students? or extending the program past the %three years? I think we need to say something here more specific, otherwise we're weakening the case ofr our proposal. } 

\newcolumntype{R}{>{\raggedleft\arraybackslash}X}%
\begin{table}
\begin{tabularx}{\textwidth} {|X|R|R|R|R|R|R|}
  \hline  
 Year &\#~New Students & \# Students Enrolled	& Student Support &Teaching Faculty Support & Postdoctoral Support & Yearly Cost\\
\hline
2015 & 0 & 0 & 0 & \$5K & 0 & \$5K\\ 
2016&	15&	15&	\$60K&	\$9K&	\$70.5K&	\$137.75K\\
2017&	15&	30&	\$120K& \$9K&	\$70.5K&	\$197.75K\\
2018&	0&	15&	\$60K&	\$9K&	\$70.5K&	\$137.75K\\
\hline
Total cost& & & \$240K & \$32K & \$211.5K &	\$483.5K\\
  \hline 
\end{tabularx}
\caption{Fully costed model for the LSST Data Science Fellowship program.}
\end{table}


\begin{table}
\begin{tabularx}{\textwidth} {|X|R|R|R|R|R|R|}
  \hline  
 Year &\#~New Students & \# Students Enrolled	& Student Support &Teaching Faculty Support & Postdoctoral Support & LSSTC Yearly Cost\\
\hline
2015 & 0 & 0 & 0 & \$5K & 0 & \$5K\\ 
2016&	15&	15&	\$60K&	\$9K&	\$0K&	\$69K\\
2017&	15&	30&	\$120K& \$9K&	\$0K&	\$129K\\
2018&	0&	15&	\$60K&	\$9K&	\$0K&	\$69K\\
\hline
LSSTC Total cost& & & \$240K & \$32K & \$0K &	\$272K\\
  \hline 
\end{tabularx}
\caption{``Northwestern'' costed model for the LSST training program. This assumes Northwestern will fund 50\% of the postdoc and the other 50\% will come from the Gates-Simonyi matching funds}
\end{table}


\begin{table}
\begin{tabularx}{\textwidth} {|X|R|R|R|R|R|R|}
  \hline  
 Year &\#~New Students & \# Students Enrolled	& Student Support &Teaching Faculty Support & Postdoctoral Support & LSSTC Yearly Cost\\
\hline
2015 & 0 & 0 & 0 & \$5K & 0 & \$5K\\ 
2016&	15&	15&	\$30K&	\$9K&	\$0K&	\$69K\\
2017&	15&	30&	\$600K& \$9K&	\$0K&	\$129K\\
2018&	0&	15&	\$30K&	\$9K&	\$0K&	\$69K\\
\hline
LSSTC Total cost& & & \$120K & \$32K & \$0K &	\$152K\\
   \hline 
\end{tabularx}
\caption{Minimal-cost model for the LSST Data Science Fellowship program. This assumes Northwestern will fund 50\% of the postdoc and the other 50\% will come from the Gates-Simonyi matching funds, while travel support for students will be provided only at 50\% of the total need.}
\end{table}


\end{document}

